Review of Edge Detectors

### 2015-03-25 edge-detection-review-3rd-revision-v3.tex
corrected MH algorithm pseudocode
v4 corrections

\documentclass{ipol}
%\ipolSetID{arced}
\ipolSetTitle{A Review of Classic Edge Detectors}
\ipolSetAuthors{Haldo Spont\'on$^1$, Juan Cardelino$^2$}
\ipolSetAffiliations{%
$^1$ IIE, UdelaR, Uruguay (\texttt{haldos@fing.edu.uy})\\
$^2$ IIE, UdelaR, Uruguay (\texttt{juanc@fing.edu.uy})}

@@

\begin{abstract}
In this paper some of the classic alternatives for edge detection in digital images are studied. The main idea 
of edge detection algorithms is to find where abrupt changes in the intensity of an image have occurred. 
The first family of algorithms reviewed in this work uses the first derivative to find the changes of intensity, 
such as Sobel, Prewitt and Roberts. In the second reviewed family, second derivatives are used, for example in algorithms 
like Marr-Hildreth and Haralick. \\
Obtained results are analyzed from a qualitative point of view (perceptual) and from a quantitative 
point of view (number of operations, execution time), considering different ways to convolve an image with a kernel (step required in some of the algorithms).
\end{abstract}

@@

\begin{ipolCode}
For all the reviewed algorithms, an open source C implementation is provided which can be downloaded from 
the IPOL publication of this article. An 
\href{\demourl}{online demonstration}%\footnote{\url{http://dev.ipol.im/~juanc/ipol_demo/sc_edges/}} 
is also available, where you can test and reproduce our results.
\end{ipolCode}

@@

% en este parrafo, puede ser cambiar trough por valley
In 1980 Marr and Hildreth \cite{segm:edge_region:marr:84:digital_step} argued that intensity changes are not independent 
of image scale, so edge detection requires the use of different size 
operators. They also argued that a sudden intensity change will be seen 
as a peak (or valley) in the first derivative or, equivalently, as a zero 
crossing in the second derivative. This algorithm is presented in Section \ref{sec:second}. 
Haralick's algorithm \cite{bb20239} is an alternative approach based on the second derivative 
which is also reviewed in Section \ref{sec:second}. This algorithm has the particularity of 
proposing a model to locally approximate the image around a point; then, using this model, 
an approximation to the second derivative of the image can be calculated analytically and 
finding edges is achieved by imposing a condition over the model parameters.

@@

Along with this paper, a detailed and well commented source code is presented, which 
implements the described algorithms. In section \ref{sec:appendix1} some common mathematical developments are presented. 
%A commented explanation of the source code can be found in section \ref{sec:appendix2}. This work also provides tools for code documentation 
%(see section \ref{sec:appendix2}). 
Results of the implemented algorithms are presented in Section \ref{sec:results}, along with examples 
to compare their performance. Conclusions are detailed in Section \ref{sec:conclusions}.

@@

\begin{algorithm}[t!]
\caption{Marr-Hildreth edge detection algorithm.}
\label{algo:mh}
\begin{algorithmic}[1]
\REQUIRE $im$ input image, standard deviation $\sigma$, kernel size $n$ and zero-crossing threshold $t_{ZC}$.
\IF{Gaussian kernel}
	\STATE Define Laplacian operator $laplacian$
	\STATE $im_{LAPL} \leftarrow$ convolution($im$,$laplacian$)
\ELSE
	\STATE $kernel \leftarrow$ generate\_kernel($n$,$\sigma$) \COMMENT{Generated Gaussian or LoG kernel}.
	\STATE $im_{SMOOTHED} \leftarrow$ convolution($im$,$kernel$)
	\STATE $im_{LAPL} \leftarrow im_{SMOOTHED}$
\ENDIF
\STATE $max_L \leftarrow 0$
\FORALL{pixel $i$ in image $im_{LAPL}$}
	\IF{$im_{LAPL}[i]>max_L$}
		\STATE $max_L \leftarrow im_{LAPL}[i]$
	\ENDIF
\ENDFOR
\FORALL{pixel $i$ in image $im_{LAPL}$, except borders}
	\FORALL{pair $(p_1,p_2)$ of opposite neighbors of $p$ in $im_{LAPL}$}
		\IF{($\mbox{sign}(im_{LAPL}[p_1])\neq\mbox{sign}(im_{LAPL}[p_2])$) \AND ($|im_{LAPL}[p_1]-im_{LAPL}[p_2]|>th_{ZC}$)}
			\STATE $im_{OUT}[i] \leftarrow 255$
		\ELSE
			\STATE $im_{OUT}[i] \leftarrow 0$
		\ENDIF
	\ENDFOR
\ENDFOR
\RETURN output image $\leftarrow im_{OUT}$
\end{algorithmic}
\end{algorithm}

>>>>

\documentclass{ipol}
%\ipolSetID{arced}
\ipolSetTitle{A Review of Classic Edge Detectors}
\ipolSetAuthors{Haldo Spont\'on\ipolAuthorMark{1}, Juan Cardelino\ipolAuthorMark{2}}
\ipolSetAffiliations{%
\ipolAuthorMark{1} Inst. Ing. Eléctrica, Universidad de la República, Uruguay (\texttt{haldos@fing.edu.uy})\\
\ipolAuthorMark{2} Inst. Ing. Eléctrica, Universidad de la República, Uruguay (\texttt{juanc@fing.edu.uy})}

@@

\begin{abstract}
In this paper some of the classic alternatives for edge detection in digital images are studied. The main idea 
of edge detection algorithms is to find where abrupt changes in the intensity of an image have occurred. 
The first family of algorithms reviewed in this work uses the first derivative to find the changes of intensity, 
such as Sobel, Prewitt and Roberts. In the second reviewed family, second derivatives are used, for example in algorithms 
like Marr-Hildreth and Haralick.
Obtained results are analyzed from a qualitative point of view (perceptual) and from a quantitative 
point of view (number of operations, execution time), considering different ways to convolve an image with a kernel (step required in some of the algorithms).
\end{abstract}

@@

\begin{ipolCode}
For all the reviewed algorithms, an open source C implementation is provided which can be downloaded from 
the IPOL publication of this article. An 
\href{\demourl}{online demonstration} is also available, where you can test and reproduce our results.
\end{ipolCode}

@@

% en este parrafo, puede ser cambiar trough por valley
In 1980 Marr and Hildreth \cite{segm:edge_region:marr:84:digital_step} argued that intensity changes are not independent 
of image scale, so edge detection requires the use of different size 
operators. They also argued that a sudden intensity change will be seen 
as a peak (or valley) in the first derivative or, equivalently, as a zero 
crossing in the second derivative. This algorithm is presented in Section \ref{sec:second}. 
Haralick's algorithm \cite{bb20239} is an alternative approach based on the second derivative 
which is also reviewed in Section \ref{sec:second}. This algorithm has the particularity of 
proposing a model to approximate the image locally around a point; then, using this model, 
an approximation to the second derivative of the image can be calculated analytically and 
finding edges is achieved by imposing a condition over the model parameters.

@@

Along with this paper, a detailed and well commented source code is presented, which 
implements the described algorithms. In section \ref{sec:appendix1} some common mathematical developments are presented. 
%A commented explanation of the source code can be found in section \ref{sec:appendix2e}. This work also provides tools for code documentation 
%(see section \ref{sec:appendix2}). 
Results of the implemented algorithms are presented in Section \ref{sec:results}, along with examples 
to compare their performance. Conclusions are detailed in Section \ref{sec:conclusions}.

@@

\begin{algorithm}[t!]
\caption{Marr-Hildreth edge detection algorithm.}
\label{algo:mh}
\begin{algorithmic}[1]
\REQUIRE $im$ input image, standard deviation $\sigma$, kernel size $n$ and zero-crossing threshold $t_{ZC}$.
\IF{Gaussian kernel}
	\STATE $kernel \leftarrow$ generate\_gaussian\_kernel($n$,$\sigma$)
	\STATE $im_{SMOOTHED} \leftarrow$ convolution($im$,$kernel$)
	\STATE Define Laplacian operator $laplacian$
	\STATE $im_{LAPL} \leftarrow$ convolution($im_{SMOOTHED}$,$laplacian$)
\ELSE
\STATE $log\_kernel \leftarrow$ generate\_log\_kernel($n$,$\sigma$)
\STATE $im_{LAPL} \leftarrow$ convolution($im$,$log\_kernel$)
\ENDIF
\STATE $max_L \leftarrow 0$
\FORALL{pixel $i$ in image $im_{LAPL}$}
	\IF{$im_{LAPL}[i]>max_L$}
		\STATE $max_L \leftarrow im_{LAPL}[i]$
	\ENDIF
\ENDFOR
\FORALL{pixel $i$ in image $im_{LAPL}$, except borders}
	\FORALL{pair $(p_1,p_2)$ of opposite neighbors of $p$ in $im_{LAPL}$}
		\IF{($\mbox{sign}(im_{LAPL}[p_1])\neq\mbox{sign}(im_{LAPL}[p_2])$) \AND ($|im_{LAPL}[p_1]-im_{LAPL}[p_2]|>th_{ZC}$)}
			\STATE $im_{OUT}[i] \leftarrow 255$
		\ELSE
			\STATE $im_{OUT}[i] \leftarrow 0$
		\ENDIF
	\ENDFOR
\ENDFOR
\RETURN output image $\leftarrow im_{OUT}$
\end{algorithmic}
\end{algorithm}
